Predicting The Success of The Female United States Artistic Gymnastics Team at The 2024 Paris Olympics Using Mean and Standard Deviation Derived Parameters

### 2024-01-09 manuscript/submission.tex
altering title, hyperref error, bib source
adding major figure changes for both men and women

\usepackage[colorlinks=true, citecolor=blue]{hyperref}

\title{UCSAS 2024 USOPC Data Challenge}
\author{Kathleen Houlihan\\
  Department of Statistics\\
  University of Connecticut
}

@@

\begin{abstract}
    For decades, gymnastics has been the most watched sport in the Summer Olympics and the United States 
    has been known for bringing gold medal-winning gymnasts to compete. In the 2020 
    Summer Olympics games in Tokyo, the United States female artistic gymnastics team took home 
    the gold medal in the Women's All Around, a bronze medal in the Women's Balance 
    Beam, a gold medal in the Women's Floor Exercise, a silver medal for the Women's
    Team, a bronze medal in the Women's Uneven Bars, and a silver medal in the Women's Vault. 
    Noting that the United States was the only country to have a team medal in all six women's artistic gymnastics 
    categories, the likely hood of the United States bringing female athletes that will medal in the 2024 
    Paris Olympics is very probable. However, in the 2020 Summer Olympic games in Tokyo, the United States 
    male artistic gymnastics team did not medal at all. For the purposes of this data challenge, I focused my 
    time on finding a statistical method that best isolates strong female gymnasts that would be prime candidates 
    for the 2024 olympics as the female USA team historically earns more medals than the male team. Later, I applied 
    my statistical method to the male team, which I anticipate will earn fewer medals than the female team, if any. 
    Noting that this data challenge can be interpreted in a number of ways, I designed my methods to maximize the 
    overall medal count while also placing greatest importance on the team all-around medal. In this paper, I will 
    identify which five female and male artistic gymnasts are best suited to be sent to the Paris 2024 
    Olympics on behalf of the United States in order to maximize the expected medal count.

@@

@@

To predict which of the United States athletes are most likely to medal on the various apparatuses, 
I will be using the cleaned data that 
is provided in the UCSAS data challenge that includes data from major domestic and international 
gymnastic competitions from 2022 and 2023. I have opted against using the data from major domestic
and international gymnastic competions from 2017 to 2021 as time, injuries, 
and other factors can have a large impact on the success of a gymnast, so the greatest predictor
of the Olympic outcomes in 2024 will be the most recent data.
\\
For the purposes of my research design, when determining the most ideal feamle olympic candidates, 
I limited the data I utilized to only include female United States
gymnasts, presuming that the United States female gymnasts have a strong chance at earning medals in any category, 
based on the results of the 2020 Summer Olympics. When determining the most ideal American male candidates, 
I did inspect the data of non United States gymnasts in order to determine which medals the United States 
male team has the greatest chance of earning. 
Furthermore, I processed the data to remove any scores of zero based on the assumption that these may 
be indicative of missing scores and that instances of these scores are rare. I further justified my decision 
to eliminate scores of zero for the purposes of knowing the true mean scores of each athlete on each 
apparatus for each successful attempt. Additionally, for some athletes and competitions, I found that the vault
apparatus was listed as "VT," "VT1," and "VT2." This discrepancy is due to the fact that athletes are permitted
two attempts on the vault apparatus. In an effort to interpret the data as accurately as possible and avoid having 
multiple statistics for one athlete on the same apparatus, I corrected this discrepancy by merging "VT," "VT1," and 
"VT2" to all be reported as "VT." Also, the data source is flawed as it alternates between reporting athlete names
in upper or lower case letters. In order to correct this, I converted all last names to lower case for consistency 
purposes. Lastly, the data source is inconsistent in the manner that it reports the first names of athletes as certain
athletes are listed using only their first name at some competitions and are listed using their first and middle names 
under the "first name" category for other competitions. In order to resolve this discrepancy, I manually verified that 
any athletes listed under two names were the same person and I further verified that there were no United States  
athletes with the same last name so that all statistical computations could be completed through grouping by last 
name alone.

@@

In order to select the appropriate athletes, I first processed the data as described in Section~\ref{sec:data}.
Then, I calculated and examined the mean and standard deviations of each American female athlete's score 
on each apparatus. I also examined the number of observations for each athlete on each apparatus to familiarize 
myself with which athletes tend to compete at the most competitions and on which apparatuses. Then, I plotted 
the mean versus the standard deviation of the ten athletes with the highest mean scores on each apparatus. 
Using this visual, I created a parameter for each apparatus that uses both mean score and standard 
deviation to identify which athletes are best suited to be considered to compete on behalf of the United 
States on each apparatus. Using this parameter, I identified the "best" five athletes that compete on
each apparatus based on both their success, a high mean score, and their inconsistency, their standard deviation. 
In cases where there was only one observation, the standard deviation value is unavailable.
For the purposes of this research, it is reasonable to ignore any athletes who have only competed once on a particular apparatus.

Next, using only the athletes I identified as the "best" on each apparatus using the parameters, 
I calculated the sum of each athletes mean score for each apparatus and the sum of each athlete's standard 
deviation for each apparatus. Similarly to before, I plotted the mean score versus the standard deviation for 
these selected athletes. This plot is intended to represent
the United States' best candidates for earning a gold medal for the individual all-around component of the Olympics.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{FemaleAthletesBB.pdf}
  \caption{Top 10 Female Candidates for the Olympic Balance Beam Apparatus}
  \label{fig:BB}
\end{figure}

\begin{table}
  \caption{Top ten female balance beam candidates and their mean minus standard deviation parameter values}
  \label{tab:tableBBP}
\centering
\begin{tabular}[t]{lccllll}
 \toprule
Last Name & Mean minus Standard Deviation\\
\midrule
biles & 14.38228\\
\midrule
mcclain & 14.13944\\
\midrule
lee & 13.68558\\
\midrule
sumanasekera & 12.94901\\
\midrule
blakely & 12.78173\\
\midrule
jones & 12.58175\\
\midrule
roberson & 12.50378\\
\midrule
wong & 12.48469\\
\midrule
dicello & 12.44901\\
\midrule
frazier & NA\\
\bottomrule
\end{tabular}
\end{table}

Based on the plot of the top ten female athletes on the balance beam apparatus I selected the parameter of 
the score standard deviation subtracted from the mean score in order to identify the best suited five 
athletes. Based on the parameter, the five "best" female athletes on balance beam are Biles, Mcclain, 
Lee, Sumanasekera, and Blakely.

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{FemaleAthletesFX.pdf}
  \caption{Top 10 Female Candidates for the Olympic Floor Exercise Apparatus}
  \label{fig:FX}
\end{figure}

\begin{table}
  \caption{Top ten female floor exercise candidates and their mean minus standard deviation parameter values}
  \label{tab:tableFXP}
\centering
\begin{tabular}[t]{lccllll}
 \toprule
Last Name & Mean minus Standard Deviation\\
\midrule
biles & 14.55017\\
\midrule
lincoln & 13.73014\\
\midrule
carey & 13.41335\\
\midrule
dicello & 13.40725\\
\midrule
chiles & 13.38463\\
\midrule
jones & 13.28955\\
\midrule
mcclain & 13.21401\\
\midrule
wong & 13.16001\\
\midrule
roberson & 12.97918\\
\midrule
frazier & NA\\
\bottomrule
\end{tabular}
\end{table}

Based on the plot of the top ten female athletes on the floor exercise apparatus I selected the parameter of 
the score standard deviation subtracted from the mean score in order to identify the best suited five 
athletes. Based on the parameter, the five "best" female athletes for floor exercise are Biles, Lincoln, Carey, 
Mcclain, Dicello, Chiles.

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{FemaleAthletesUB.pdf}
  \caption{Top 10 Female Candidates for the Olympic Uneven Bar Apparatus}
  \label{fig:UB}
\end{figure}

\begin{table}
  \caption{Top 10 female uneven bars candidates and their mean minus half their standard deviation parameter values}
  \label{tab:tableUBP}
\centering
\begin{tabular}[t]{lccllll}
 \toprule
Last Name & Mean minus (0.5)Standard Deviation\\
\midrule
jones & 14.26300\\
\midrule
biles & 14.16786\\
\midrule
wong & 13.58939\\
\midrule
matthews & 13.50008\\
\midrule
blakely & 13.48821\\
\midrule
greaves & 13.47197\\
\midrule
miller & 13.37735\\
\midrule
chiles & 13.33873\\
\midrule
zhou & 13.27921\\
\midrule
walker & NA\\
\bottomrule
\end{tabular}
\end{table}

Based on the plot of the top ten female athletes on the uneven bars apparatus I selected the parameter of 
half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
athletes. Based on the parameter, the five "best" female athletes for uneven bars are Jones, Biles, Wong, 
Matthews, and Blakely.

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{FemaleAthletesVT.pdf}
  \caption{Top 10 Female Candidates for the Olympic Vault Apparatus}
  \label{fig:VT}
\end{figure}

\begin{table}
  \caption{Top ten female vault candidates and their mean minus half their standard deviation parameter values}
  \label{tab:tableVTP}
\centering
\begin{tabular}[t]{lccllll}
 \toprule
Last Name & Mean minus (0.5)Standard Deviation\\
\midrule
biles & 14.77708\\
\midrule
carey & 14.35999\\
\midrule
mcclain & 14.20297\\
\midrule
jones & 14.19854\\
\midrule
chiles & 14.10796\\
\midrule
sumanasekera & 14.06650\\
\midrule
blakely & 14.05000\\
\midrule
dicello & 13.88174\\
\midrule
richardson & NA\\
\midrule
torry & NA\\
\bottomrule
\end{tabular}
\end{table}

Based on the plot of the top ten athletes on the vault apparatus I selected the parameter of 
half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
athletes. Based on the parameter, the five "best" female athletes on vault are Biles, Carey, Mcclain, Jones, 
and Chiles.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{FemaleAthletesAACandidates.pdf}
  \caption{The best candidates for the female individual all-around medal}
  \label{fig:IAA}
\end{figure}

\begin{table}
  \caption{Summed means and standard deviations for the selected female individual all-around candidates}
  \label{tab:tableIAA}
\centering
\begin{tabular}[t]{lccllll}
 \toprule
Last Name & Summed Means & Summed Standard Deviations\\
\midrule
biles & 58.70640 & 1.124194\\
\midrule
jones & 55.82380 & 1.826863\\
\midrule
mcclain & 55.68333 & 1.741128\\
\midrule
blakely & 54.52908 & 2.017384\\
\midrule
chiles & 54.50832 & 1.941718\\
\midrule
carey & 54.38711 & 1.770006\\
\midrule
wong & 54.28354 & 1.936921\\
\midrule
dicello & 54.11984 & 1.958245\\
\midrule
sumanasekera & 53.15505 & 1.589704\\
\midrule
lincoln & 52.39283 & 3.048907\\
\midrule
matthews & 52.33400 & 1.556188\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{FemaleAthletesAA3.pdf}
  \caption{Top 10 combinations of three female candidates for the team all-around medal}
  \label{fig:AA3}
\end{figure}

\begin{table}
  \caption{The five combinations of three United States female athletes with the highest combined mean scores}
  \label{tab:tableAA3}
\centering
\begin{tabular}[t]{lccllll}
 \toprule
Combination & Combined Score & Standard Deviation Sum\\
\midrule
biles, jones, mcclain & 170.2135 & 4.692185\\
\midrule
biles, blakely, jones & 169.0593 & 4.968440\\
\midrule
biles, chiles, jones & 169.0385 & 4.892774\\
\midrule
biles, blakely, mcclain & 168.9188 & 4.882706\\
\midrule
biles, carey, jones
& 168.9173 & 4.721063
\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{FemaleAthletesAA4.pdf}
  \caption{Top 10 combinations of four female candidates for the team all-around medal}
  \label{fig:AA4}
\end{figure}

\begin{table}
  \caption{The five combinations of four United States female athletes with the highest combined mean scores}
  \label{tab:tableAA4}
\centering
\begin{tabular}[t]{lccllll}
 \toprule
Combination & Combined Score & Standard Deviation Sum\\
\midrule
biles, blakely, jones, mcclain & 224.7426 & 6.709569\\
\midrule
biles, chiles, jones, mcclain & 224.7218 & 6.633903\\
\midrule
biles, carey, jones, mcclain & 224.6006 & 6.462191\\
\midrule
biles, jones, mcclain, wong & 224.4971 & 6.629105\\
\midrule
biles, dicello, jones, mcclain & 224.3334 & 6.650430\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{MaleAthletesFX.pdf}
    \caption{Top 10 Male Candidates for the Olympic Floor Exercise Apparatus}
    \label{fig:FXM}
  \end{figure}

  Based on the plot of the top ten male athletes on the floor exercise apparatus I selected the parameter of 
  the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five "best" male athletes on floor exercise are Juda, Mccool, Young, Moldauer, 
  and Hong.

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{MaleAthletesVT.pdf}
    \caption{Top 10 Male Candidates for the Olympic Vault Apparatus}
    \label{fig:VTM}
  \end{figure}

  Based on the plot of the top ten male athletes on the vault apparatus I selected the parameter of 
  the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five "best" male athletes on vault are Hale, Young, Malone, Briones, and 
  Juda.

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{MaleAthletesSR.pdf}
    \caption{Top 10 Male Candidates for the Olympic Still Rings Apparatus}
    \label{fig:SRM}
  \end{figure}

  Based on the plot of the top ten male athletes on the still rings apparatus I selected the parameter of 
  half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five "best" male athletes on still rings are Whittenburg, Briones, Diab, 
  Hong, and Loos.

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{MaleAthletesPH.pdf}
    \caption{Top 10 Male Candidates for the Olympic Pommel Horse Apparatus}
    \label{fig:PHM}
  \end{figure}

  Based on the plot of the top ten male athletes on the pommel horse apparatus I selected the parameter of 
  half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five "best" male athletes on pommel horse are Nedoroscik, Young, Malone, 
  Dang, and Bock.

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{MaleAthletesPB.pdf}
    \caption{Top 10 Male Candidates for the Olympic Parallel Bars Apparatus}
    \label{fig:PBM}
  \end{figure}

  Based on the plot of the top ten male athletes on the parallel bars apparatus I selected the parameter of 
  half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five "best" male athletes on parallel bars are Hong, Chou, Walker, Malone, 
  and Phillips.

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{MaleAthletesHB.pdf}
    \caption{Top 10 Male Candidates for the Olympic High Bar Apparatus}
    \label{fig:HBM}
  \end{figure}

  Based on the plot of the top ten male athletes on the high bar apparatus I selected the parameter of 
  half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five "best" male athletes on high bar are Malone, Juda, Briones, Bold, and 
  Nguyen.

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{MaleAthletesAACandidates.pdf}
    \caption{The best candidates for the male individual all-around medal}
    \label{fig:IAAM}
  \end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{MaleAthletesAA3.pdf}
    \caption{Top 10 combinations of three male candidates for the team all-around medal}
    \label{fig:MAA3}
  \end{figure}

@@

  \begin{figure}
    \centering
    \includegraphics[scale=0.6]{MaleAthletesAA4.pdf}
    \caption{Top 10 combinations of four male candidates for the team all-around medal}
    \label{fig:MAA4}
  \end{figure}

@@

For the floor exercise apparatus, I selected the parameter: mean score minus standard deviation. This parameter 
is reasonable as well as it selects athletes with high means and low standard deviations first (Biles then Lincoln) 
and selects athletes with lower standard deviations when means are similar (Carey, Dicello, and Chiles). Note, the 
order in which Carey, Dicello, and Chiles "should" be selected is debatable, but it is most important that the 
parameter selected these athletes as opposed to athletes that have similar mean scores but higher standard deviations 
(Jones and Mcclain) or low standard deviations and low low mean scores (Wong).

@@

Similarly, I used the same method to determine the "best" male all-around candidates as I did for females. From 
Figure~\ref{fig:IAAM} I determined that Juda and Malone are the best male Individual all-around candidates. From 
Figure~\ref{fig:MAA3} I determined that Juda, Malone, and Young are the best male team all-around finalists 
candidates. Additionally, from Figure~\ref{fig:MAA4} I determined that Moldauer is the best candidate to also compete 
with Juda, Malone, and Young in the all-around qualifying round. Comparatively, determining the "best" fifth Olympic 
teammate is challenging as there are several apparatuses where our "best" competitor is not among the four 
previously selected teammates. However, through examining the the mean score for every male gymnast on every apparatus 
we can see that many of our "best" candidates are still ranked very poorly globally, and therefore would not have a 
strong liklihood of medalling at the olympics. Among all of the male USA athletes that are not yet selected, the only 
athlete that has a relatively decent chance of medaling is Hale, who ranks 6th globally for the Vault apparatus.

@@

>>>>

\title{Predicting The Success of The Female United States Artistic Gymnastics Team at The 2024 Paris Olympics 
Using Mean and Standard Deviation Derived Parameters}
\author{Kathleen Houlihan\\
  Department of Statistics\\
  University of Connecticut
}

@@

\begin{abstract}

@@

\section{Introduction}
\label{sec:intro}

For decades, gymnastics has been the most watched sport in the Summer Olympics and the United States 
has been known for bringing gold medal-winning gymnasts to compete. In the 2020 
Summer Olympics games in Tokyo, the United States female artistic gymnastics team took home 
the gold medal in the Women's All Around, a bronze medal in the Women's Balance 
Beam, a gold medal in the Women's Floor Exercise, a silver medal for the Women's
Team, a bronze medal in the Women's Uneven Bars, and a silver medal in the Women's Vault. 
Noting that the United States was the only country to have a team medal in all six women's artistic gymnastics 
categories, the likely hood of the United States bringing female athletes that will medal in the 2024 
Paris Olympics is very probable. However, in the 2020 Summer Olympic games in Tokyo, the United States 
male artistic gymnastics team did not medal at all. For the purposes of this data challenge, I focused my 
time on finding a statistical method that best isolates strong female gymnasts that would be prime candidates 
for the 2024 olympics as the female USA team historically earns more medals than the male team. Later, I applied 
my statistical method to the male team, which I anticipate will earn fewer medals than the female team, if any. 
Noting that this data challenge can be interpreted in a number of ways, I designed my methods to maximize the 
overall medal count while also placing greatest importance on the team all-around medal. In this paper, I will 
identify which five female and male artistic gymnasts are best suited to be sent to the Paris 2024 
Olympics on behalf of the United States in order to maximize the expected medal count.

Currently there is no literature available relevant to predicting gymnastics outcomes at the 2024 
Paris Summer Olympics or at recent olympics. However, there is literature that relates to various aspects of 
the data that will be used to form my predictions. For example, the article 'The Prediction of All-Around 
Event Final Score Based On D and E Score Factors In Women's Artistic Gymnastics' aims to determine the 
impact of individual apparatus difficulty and execution scores on overall score. 
Expertise from this article could be useful in allowing us to understand how D and E scores, 
which are available in my data, play an effort in predicting the overall score. Additionally, the paper 
'Performance rating in men’s world elite artistic gymnastics: a status-quo study on scoring 
tendencies at Olympic Games following rule changes' also can play a similar role in predicting overall 
score. This paper's objective is to analyze the effect on elite artistic gymnastics had by the 2006 rule change 
which aimed to reduce the dominance of difficulty while strengthening performance differentiation in 
competition. Notably, the paper found that the rule change resulted in significantly enhanced performance 
differentiation with the execution or E score being the main predictor of the final score as opposed to 
the difficulty or D score. Furthermore, in the article 'How Apparatus Difficulty Scores Affect All 
Around Results in Men's Artistic Gymnastics,' the authors use regression, cluster, and ANOVA analysis to 
analyze how difficulty scores predict overall score differently across different apparatuses. However, in this 
article the authors found that while difficulty scores may have predictive tendancies for some apparatuses, 
the strength of the prediction varied drastically from apparatus to apparatus, with there being little to no 
predictive ability for some apparatuses. Ultimately, the available literature on predicting artistic 
gymnastic scores seemed generally limited to how we can use the E and D scores provided in the data sets 
to predict overall score, but some aspects of this literature casted doubt on whether E and D scores should be 
used as predictors at all.

@@

To predict which of the United States athletes are most likely to medal on the various apparatuses, 
I will be using the cleaned data that 
is provided in the UCSAS data challenge that includes data from major domestic and international 
gymnastic competitions from 2022 and 2023. I have opted against using the data from major domestic
and international gymnastic competions from 2017 to 2021 as time, injuries, 
and other factors can have a large impact on the success of a gymnast, so the greatest predictor
of the Olympic outcomes in 2024 will be the most recent data.
\\
For the purposes of my research design, when determining the most ideal feamle olympic candidates, 
I limited the data I utilized to only include female United States
gymnasts, presuming that the United States female gymnasts have a strong chance at earning medals in any category, 
based on the results of the 2020 Summer Olympics. When determining the most ideal American male candidates, 
I did inspect the data of non United States gymnasts in order to determine which medals the United States 
male team has the greatest chance of earning. 
Furthermore, I processed the data to remove any scores of zero based on the assumption that these may 
be indicative of missing scores and that instances of these scores are rare. I further justified my decision 
to eliminate scores of zero for the purposes of knowing the true mean scores of each athlete on each 
apparatus for each successful attempt. Additionally, for some athletes and competitions, I found that the vault
apparatus was listed as 'VT,' 'VT1,' and 'VT2.' This discrepancy is due to the fact that athletes are permitted
two attempts on the vault apparatus. In an effort to interpret the data as accurately as possible and avoid having 
multiple statistics for one athlete on the same apparatus, I corrected this discrepancy by merging 'VT,' 'VT1,' and 
'VT2' to all be reported as 'VT.' Also, the data source is flawed as it alternates between reporting athlete names
in upper or lower case letters. In order to correct this, I converted all last names to lower case for consistency 
purposes. Lastly, the data source is inconsistent in the manner that it reports the first names of athletes as certain
athletes are listed using only their first name at some competitions and are listed using their first and middle names 
under the 'first name' category for other competitions. In order to resolve this discrepancy, I manually verified that 
any athletes listed under two names were the same person and I further verified that there were no United States  
athletes with the same last name so that all statistical computations could be completed through grouping by last 
name alone.

@@

In order to select the appropriate athletes, I first processed the data as described in Section~\ref{sec:data}.
Then, I calculated and examined the mean and standard deviations of each American female athlete's score 
on each apparatus. I also examined the number of observations for each athlete on each apparatus to familiarize 
myself with which athletes tend to compete at the most competitions and on which apparatuses. Then, I plotted 
the mean versus the standard deviation of the ten athletes with the highest mean scores on each apparatus. 
Using this visual, I created a parameter for each apparatus that uses both mean score and standard 
deviation to identify which athletes are best suited to be considered to compete on behalf of the United 
States on each apparatus. Using this parameter, I identified the 'best' five athletes that compete on
each apparatus based on both their success, a high mean score, and their inconsistency, their standard deviation. 
In cases where there was only one observation, the standard deviation value is unavailable.
For the purposes of this research, it is reasonable to ignore any athletes who have only competed once on a particular apparatus.

Next, using only the athletes I identified as the 'best' on each apparatus using the parameters, 
I calculated the sum of each athletes mean score for each apparatus and the sum of each athlete's standard 
deviation for each apparatus. Similarly to before, I plotted the mean score versus the standard deviation for 
these selected athletes. This plot is intended to represent
the United States' best candidates for earning a gold medal for the individual all-around component of the Olympics.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.7]{FemaleAthleteApparatus.pdf}
  \caption{Top 10 Female Candidates for the Olympic Balance Beam, Floor Exercise, Uneven Bars, and Vault Apparatus}
  \label{fig:FA}
\end{figure}

\begin{table}
  \caption{Top ten female candidates for each apparatus based on mean score alone and 
  their $\mu - \alpha \sigma$ parameter values}
  \label{tab:tableBBP}
\centering
\begin{tabular}[t]{llllllll}
 \toprule
  \multicolumn{2}{c}{BB} & \multicolumn{2}{c}{FX} & \multicolumn{2}{c}{UB} & \multicolumn{2}{c}{VT}\\
  \cmidrule(lr){1-2}\cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}
Athlete & $\mu - \sigma$ & Athlete & $\mu - \sigma$ & Athlete & $\mu - 0.5 \sigma$ & Athlete & $\mu - 0.5 \sigma$\\
\midrule
biles & 14.38 & biles & 14.55 & jones & 14.26 & Biles & 14.78\\
\midrule
mcclain & 14.14 & lincoln & 13.73 & biles & 14.17 & Carey & 14.36\\
\midrule
lee & 13.69 & carey & 13.41 & wong & 13.59 & Mcclain & 14.20\\
\midrule
sumanasekera & 12.95 & dicello & 13.4 & matthews & 13.50 & Jones & 14.20\\
\midrule
blakely & 12.78 & chiles & 13.38 & blakely & 13.49 & Chiles & 14.11\\
\midrule
jones & 12.58 & jones & 13.29 & greaves & 13.47 & sumanasekera & 14.07\\
\midrule
roberson & 12.50 & mcclain & 13.21 & miller & 13.38 & blakely & 14.05\\
\midrule
wong & 12.48 & wong & 13.16 & chiles & 13.34 & dicello & 13.88\\
\midrule
dicello & 12.45 & roberson & 12.98 & zhou & 13.28 & richardson & NA\\
\midrule
frazier & NA & frazier & NA & walker & NA & torry & NA\\
\bottomrule
\end{tabular}
\end{table}

Based on the plot of the top ten female athletes on the balance beam apparatus I selected the parameter of 
the score standard deviation subtracted from the mean score in order to identify the best suited five 
athletes. Based on the parameter, the five 'best' female athletes on balance beam are Biles, Mcclain, 
Lee, Sumanasekera, and Blakely.

Based on the plot of the top ten female athletes on the floor exercise apparatus I selected the parameter of 
the score standard deviation subtracted from the mean score in order to identify the best suited five 
athletes. Based on the parameter, the five 'best' female athletes for floor exercise are Biles, Lincoln, Carey, 
Mcclain, Dicello, Chiles.

Based on the plot of the top ten female athletes on the uneven bars apparatus I selected the parameter of 
half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
athletes. Based on the parameter, the five 'best' female athletes for uneven bars are Jones, Biles, Wong, 
Matthews, and Blakely.

Based on the plot of the top ten athletes on the vault apparatus I selected the parameter of 
half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
athletes. Based on the parameter, the five 'best' female athletes on vault are Biles, Carey, Mcclain, Jones, 
and Chiles.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.55]{FemaleMultipleAllAround.pdf}
  \caption{The best candidates for the female individual all-around and team all around medals}
  \label{fig:FAA}
\end{figure}

@@

@@

@@

  Based on the plot of the top ten male athletes on the floor exercise apparatus I selected the parameter of 
  the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five 'best' male athletes on floor exercise are Juda, Mccool, Young, Moldauer, 
  and Hong.

  Based on the plot of the top ten male athletes on the vault apparatus I selected the parameter of 
  the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five 'best' male athletes on vault are Hale, Young, Malone, Briones, and 
  Juda.

  Based on the plot of the top ten male athletes on the still rings apparatus I selected the parameter of 
  half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five 'best' male athletes on still rings are Whittenburg, Briones, Diab, 
  Hong, and Loos.

  Based on the plot of the top ten male athletes on the pommel horse apparatus I selected the parameter of 
  half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five 'best' male athletes on pommel horse are Nedoroscik, Young, Malone, 
  Dang, and Bock.

  Based on the plot of the top ten male athletes on the parallel bars apparatus I selected the parameter of 
  half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five 'best' male athletes on parallel bars are Hong, Chou, Walker, Malone, 
  and Phillips.

  Based on the plot of the top ten male athletes on the high bar apparatus I selected the parameter of 
  half of the score standard deviation subtracted from the mean score in order to identify the best suited five 
  athletes. Based on the parameter, the five 'best' male athletes on high bar are Malone, Juda, Briones, Bold, and 
  Nguyen.

@@

@@

@@

@@

For the floor exercise apparatus, I selected the parameter: mean score minus standard deviation. This parameter 
is reasonable as well as it selects athletes with high means and low standard deviations first (Biles then Lincoln) 
and selects athletes with lower standard deviations when means are similar (Carey, Dicello, and Chiles). Note, the 
order in which Carey, Dicello, and Chiles 'should' be selected is debatable, but it is most important that the 
parameter selected these athletes as opposed to athletes that have similar mean scores but higher standard deviations 
(Jones and Mcclain) or low standard deviations and low low mean scores (Wong).

@@

\appendix

\section{Appendix: Application of the Mean and Standard Deviation Derived Parameters Methods For The Male 
United States Artistic Gymnastics Team at The 2024 Paris Olympics}

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{MaleAthleteApparatus.pdf}
  \caption{Top 10 Male Candidates for the Olympic Floor Exercise, Pommel Horse, Still Rings, Vault, Parallel Bars, 
  and High Bar Apparatus}
  \label{fig:FA}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=0.7]{MaleMultipleAllAround.pdf}
  \caption{The best candidates for the male individual all-around and team all around medals}
  \label{fig:FAA}
\end{figure}

Similar to the female model, I used the same method to determine the 'best' male all-around candidates as I did for females. From 
Figure~\ref{fig:IAAM} I determined that Juda and Malone are the best male Individual all-around candidates. From 
Figure~\ref{fig:MAA3} I determined that Juda, Malone, and Young are the best male team all-around finalists 
candidates. Additionally, from Figure~\ref{fig:MAA4} I determined that Moldauer is the best candidate to also compete 
with Juda, Malone, and Young in the all-around qualifying round. Comparatively, determining the 'best' fifth Olympic 
teammate is challenging as there are several apparatuses where our 'best' competitor is not among the four 
previously selected teammates. However, through examining the the mean score for every male gymnast on every apparatus 
we can see that many of our 'best' candidates are still ranked very poorly globally, and therefore would not have a 
strong liklihood of medalling at the olympics. Among all of the male USA athletes that are not yet selected, the only 
athlete that has a relatively decent chance of medaling is Hale, who ranks 6th globally for the Vault apparatus.

@@

\bibliographystyle{chicago}
\bibliography{citations}
